---
title: "Paper_clean"
author: "Cristiano parmeggiani"
date: "2025-11-21"
output: html_document
---

```{r setup, include=FALSE
  knitr::opts_chunk$set(echo = TRUE)}

#The purpose of this file is to examin the relationship between two factors of placentation, interdigitation and invasiveness, and life history vairaibles, gestation length, neonate body size, litter size, and interbirth interval. Each of these variables are chosen for their importance in resourse exchange during pregnancy, following the prediction that with an increased intrusiveness of placental interdigitation or invasiveness, which allows for more resource exchange between fetus and gestator, there will be compensation by reducing the intensitiy of the life history variable in question (as in, shorter gestation length, etc.). A fuller version of introduction and conclusions can be found in the paper.

```

```{r}

# Libraries 

library(tidyverse) #basic data analysis packages
library(janitor)
library(ggfortify)
library(ggpmisc)
library(ggplot2)
library(corrplot)
library(GGally)#graphic packages
library(paletteer)
library(patchwork)
library(ape)#next 4 are for phylogenic correction
library(geiger)
library(nlme)
library(phytools)
library(ggtree)
library(caper)
library(treeio)
library(tree)
library(ggrepel)
library(geiger)
library(ggnewscale)
library(phytools)
library(stringdist)
library(ggthemes)
library(multcomp)
library(broom)
library(scales) 
library(emmeans)
library(purrr)
library(tidyr)
library(knitr)
library(kableExtra)
library(gridExtra)
library(grid)
library(viridis)
library(tibble)
library(openxlsx)
library(rmarkdown)
library(dplyr)

```

```{r}

#First step in the analysis is to visualize the placenta data on a phylogeny.

#### Phylogeny 


#Import Newick tree
mammal_phy <- ape::read.tree("mammal_tree.nwk")
mammal_phy$tip.label <- tolower(mammal_phy$tip.label)

#Import datasets
placenta <- readr::read_csv("placenta.csv")
gestation_length <- readr::read_csv("gestation_length.csv")
neonate_body_size <- readr::read_csv("neonate_body_size.csv")
litter_size <- readr::read_csv("litter_size.csv")
interbirth_interval <- readr::read_csv("interbirth_interval.csv")

placenta$species <- tolower(placenta$species)

gestree <- prune_tree(gestation_length)
gestree_female <- prune_tree(gestation_length_female)
litsizetree <- prune_tree(litsize)
litsizetree_female <- prune_tree(litsize_female)
littree <- prune_tree(relative_litter_mass)
littree_female <- prune_tree(relative_litter_mass_female)
neotree <- prune_tree(neonate_body_size)
neotree_female <- prune_tree(neonate_body_size_female)
intervaltree <- prune_tree(interbirth_interval)
intervaltree_female <- prune_tree(interbirth_interval_female)

```

```{r}

# 06 — Comparative data and SE preparation -----------------------------------

make_comp_data <- function(data, tree) {
  data <- data %>%
    filter(species %in% tree$tip.label) %>%
    distinct(species, .keep_all = TRUE) %>%
    as.data.frame()
  rownames(data) <- NULL

  comp <- comparative.data(
    phy = tree,
    data = data,
    names.col = "species",
    vcv = FALSE,
    vcv.dim = 2,
    na.omit = FALSE,
    force.root = FALSE,
    warn.dropped = TRUE
  )

  if (!"species" %in% colnames(comp$data)) {
    comp$data$species <- rownames(comp$data)
  }

  comp$data <- comp$data %>%
    mutate(
      invasiveness = factor(invasiveness,
        levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial")
      ),
      interdigitation = factor(interdigitation,
        levels = c("Villous", "Trabecular", "Labyrinthine")
      )
    )

  comp
}

comp_phylo_ges <- make_comp_data(gestation_length, gestree)
comp_phylo_ges_female <- make_comp_data(gestation_length_female, gestree_female)
comp_phylo_litsize <- make_comp_data(litsize, litsizetree)
comp_phylo_litsize_female <- make_comp_data(litsize_female, litsizetree_female)
comp_phylo_lit <- make_comp_data(relative_litter_mass, littree)
comp_phylo_lit_female <- make_comp_data(relative_litter_mass_female, littree_female)
comp_phylo_neo <- make_comp_data(neonate_body_size, neotree)
comp_phylo_neo_female <- make_comp_data(neonate_body_size_female, neotree_female)
comp_phylo_interval <- make_comp_data(interbirth_interval, intervaltree)
comp_phylo_interval_female <- make_comp_data(interbirth_interval_female, intervaltree_female)

prepare_model_and_se <- function(response_var,
                                 predictor_vars,
                                 data_df,
                                 se_source_df,
                                 phylo_obj,
                                 log_response = TRUE,
                                 default_se = 0.01) {
  
  # --- ensure species column & standardize casing ---
  if (!"species" %in% names(data_df)) {
    if ("spp" %in% names(data_df)) {
      data_df$species <- data_df$spp
    } else {
      stop("`species` (or `spp`) column not found in data_df.")
    }
  }
  
  data_df$species       <- tolower(data_df$species)
  se_source_df$species  <- tolower(se_source_df$species)
  phylo_obj$tip.label   <- tolower(phylo_obj$tip.label)
  
  # --- keep only species present in the tree ---
  data_df <- dplyr::filter(data_df, species %in% phylo_obj$tip.label)
  
  # --- build per-species SE on the *linear* scale from the source df ---
  se_df <- se_source_df %>%
    dplyr::filter(species %in% data_df$species) %>%
    dplyr::group_by(species) %>%
    dplyr::summarise(
      n_obs     = dplyr::n(),
      sd_resp   = stats::sd(.data[[response_var]], na.rm = TRUE),
      se_linear = dplyr::if_else(n_obs > 1, sd_resp / sqrt(n_obs), NA_real_),
      .groups = "drop"
    )
  
  # --- assign default SE for single-observation species ---
  se_df$se_linear[is.na(se_df$se_linear)] <- default_se
  
  # --- join traits + predictors + SEs ---
  df <- data_df %>%
    dplyr::select(species, dplyr::all_of(c(response_var, predictor_vars))) %>%
    dplyr::left_join(se_df[, c("species", "se_linear")], by = "species")
  
  # --- if modeling log(response), delta-method SE = SE_linear / response ---
  if (log_response) {
    df <- dplyr::filter(df, .data[[response_var]] > 0) # required for logs and delta method
    df$se_use <- df$se_linear / df[[response_var]]
  } else {
    df$se_use <- df$se_linear
  }
  
  # --- make predictors factors if they are categorical ---
  for (v in predictor_vars) {
    if (is.character(df[[v]])) df[[v]] <- factor(df[[v]])
  }
  
  # --- drop any rows with NA in response/predictors/SE ---
  df <- tidyr::drop_na(df, dplyr::all_of(c(response_var, predictor_vars, "se_use")))
  
  # --- order by tree tip order and keep spp copy ---
  df <- df[match(phylo_obj$tip.label, df$species), , drop = FALSE]
  df <- df[stats::complete.cases(df$species), , drop = FALSE]
  df$spp <- df$species
  
  # --- build a model.frame with na.fail to catch anything early ---
  form <- stats::as.formula(
    paste(response_var, "~", paste(c(predictor_vars, "spp"), collapse = " + "))
  )
  mf <- stats::model.frame(form, data = df, na.action = stats::na.fail)
  
  # --- align SE vector to the model.frame rows ---
  se_out <- df$se_use[match(mf$spp, df$species)]
  names(se_out) <- mf$spp
  
  # --- final safety checks ---
  if (anyNA(se_out) || any(!is.finite(se_out)))
    stop("SE vector contains NA/Inf after alignment & delta method. Check zeros/missing in response or se_source_df.")
  
  list(
    model_frame = mf,
    se          = se_out
  )
}

```

```{r}

# --- Use trait-specific datasets & matching phylogenetic trees ---

trait_datasets <- list(
  gestation_length = gestation_length,
  litter_size = litsize,
  relative_litter_mass = relative_litter_mass,
  neonate_over_adult = neonate_body_size,
  interbirth_interval = interbirth_interval
)

trait_trees <- list(
  gestation_length = gestree,
  litter_size = litsizetree,
  relative_litter_mass = littree,
  neonate_over_adult = neotree,
  interbirth_interval = intervaltree
)

responses <- names(trait_datasets)
factors   <- c("interdigitation", "invasiveness")
control_vars <- c("longevity", "body_mass_g")
results <- list()

for (resp in responses) {
  df  <- trait_datasets[[resp]]
  phy <- trait_trees[[resp]]
  
  for (fac in factors) {
    res <- prepare_model_and_se(
      response_var   = resp,
      predictor_vars = c(fac, control_vars),
      data_df        = df,
      se_source_df   = df,
      phylo_obj      = phy,
      log_response   = TRUE
    )
    
    results[[paste(resp, fac, sep = "_")]] <- res
  }
}

female_trait_datasets <- list(
  gestation_length = gestation_length_female,
  litter_size = litsize_female,
  relative_litter_mass_female = relative_litter_mass_female,
  neonate_over_adult = neonate_body_size_female,
  interbirth_interval = interbirth_interval_female
)

female_trait_trees <- list(
  gestation_length = gestree_female,
  litter_size = litsizetree_female,
  relative_litter_mass_female = littree_female,
  neonate_over_adult = neotree_female,
  interbirth_interval = intervaltree_female
)

female_control_vars <- c("longevity", "female_body_mass_g")
female_results <- list()

female_responses <- names(female_trait_datasets)


for (resp in female_responses) {
  df  <- female_trait_datasets[[resp]]
  phy <- female_trait_trees[[resp]]
  
  for (fac in factors) {
    res <- prepare_model_and_se(
      response_var   = resp,
      predictor_vars = c(fac, female_control_vars),
      data_df        = df,
      se_source_df   = df,
      phylo_obj      = phy,
      log_response   = TRUE
    )
    
    female_results[[paste(resp, fac, sep = "_")]] <- res
  }
}

```

```{r}

# SE can now be accessed.

# Gestation length
model_inter_ges <- results[["gestation_length_interdigitation"]]$model_frame
se_inter_ges    <- results[["gestation_length_interdigitation"]]$se

model_inv_ges <- results[["gestation_length_invasiveness"]]$model_frame
se_inv_ges    <- results[["gestation_length_invasiveness"]]$se

model_inter_ges_female <- female_results[["gestation_length_interdigitation"]]$model_frame
se_inter_ges_female    <- female_results[["gestation_length_interdigitation"]]$se

model_inv_ges_female <- female_results[["gestation_length_invasiveness"]]$model_frame
se_inv_ges_female    <- female_results[["gestation_length_invasiveness"]]$se

# Litter size
model_inter_litsize <- results[["litter_size_interdigitation"]]$model_frame
se_inter_litsize    <- results[["litter_size_interdigitation"]]$se

model_inv_litsize <- results[["litter_size_invasiveness"]]$model_frame
se_inv_litsize    <- results[["litter_size_invasiveness"]]$se

model_inter_litsize_female <- female_results[["litter_size_interdigitation"]]$model_frame
se_inter_litsize_female    <- female_results[["litter_size_interdigitation"]]$se

model_inv_litsize_female <- female_results[["litter_size_invasiveness"]]$model_frame
se_inv_litsize_female    <- female_results[["litter_size_invasiveness"]]$se

# Litter mass
model_inter_litter <- results[["relative_litter_mass_interdigitation"]]$model_frame
se_inter_litter    <- results[["relative_litter_mass_interdigitation"]]$se

model_inv_litter <- results[["relative_litter_mass_invasiveness"]]$model_frame
se_inv_litter    <- results[["relative_litter_mass_invasiveness"]]$se

model_inter_litter_female <- female_results[["relative_litter_mass_female_interdigitation"]]$model_frame
se_inter_litter_female    <- female_results[["relative_litter_mass_female_interdigitation"]]$se

model_inv_litter_female <- female_results[["relative_litter_mass_female_invasiveness"]]$model_frame
se_inv_litter_female    <- female_results[["relative_litter_mass_female_invasiveness"]]$se

# Neonate body weight / adult body weight
model_inter_neo <- results[["neonate_over_adult_interdigitation"]]$model_frame
se_inter_neo    <- results[["neonate_over_adult_interdigitation"]]$se

model_inv_neo <- results[["neonate_over_adult_invasiveness"]]$model_frame
se_inv_neo    <- results[["neonate_over_adult_invasiveness"]]$se

model_inter_neo_female <- female_results[["neonate_over_adult_interdigitation"]]$model_frame
se_inter_neo_female    <- female_results[["neonate_over_adult_interdigitation"]]$se

model_inv_neo_female <- female_results[["neonate_over_adult_invasiveness"]]$model_frame
se_inv_neo_female    <- female_results[["neonate_over_adult_invasiveness"]]$se

# Interbirth interval
model_inter_interval <- results[["interbirth_interval_interdigitation"]]$model_frame
se_inter_interval    <- results[["interbirth_interval_interdigitation"]]$se

model_inv_interval <- results[["interbirth_interval_invasiveness"]]$model_frame
se_inv_interval    <- results[["interbirth_interval_invasiveness"]]$se

model_inter_interval_female <- female_results[["interbirth_interval_interdigitation"]]$model_frame
se_inter_interval_female    <- female_results[["interbirth_interval_interdigitation"]]$se

model_inv_interval_female <- female_results[["interbirth_interval_invasiveness"]]$model_frame
se_inv_interval_female    <- female_results[["interbirth_interval_invasiveness"]]$se

```

```{r PGLS function}


library(nlme)
library(rms)
library(phytools)
library(tidyverse)
library(geiger)
library(ape)
library(caper)
library(stringr)

modPgls.SEy = function (model, data, corClass = corBrownian, tree, se = NULL, 
                        method = c("REML", "ML"), interval = c(0, 1000), corClassValue=1, sig2e=NULL, ...) 
{
  Call <- match.call()
  corfunc <- corClass
  spp <- rownames(data)
  data <- cbind(data, spp)
  if (is.null(se)) 
    se <- setNames(rep(0, Ntip(tree)), tree$tip.label)[spp]
  else se <- se[spp]
  
  lk <- function(sig2e, data, tree, model, ve, corfunc, spp) {
    tree$edge.length <- tree$edge.length * sig2e
    ii <- sapply(1:Ntip(tree), function(x, e) which(e == 
                                                      x), e = tree$edge[, 2])
    tree$edge.length[ii] <- tree$edge.length[ii] + ve[tree$tip.label]
    vf <- diag(vcv(tree))[spp]
    w <- varFixed(~vf)
    COR <- corfunc(corClassValue, tree, form = ~spp, ...)
    fit <- gls(model, data = cbind(data, vf), correlation = COR, 
               method = method, weights = w)
    -logLik(fit)
  }
  
  if (is.null(sig2e)) {
    fit <- optimize(lk, interval = interval, data = data, tree = tree, 
                    model = model, ve = se^2, corfunc = corfunc, spp = spp)
    sig2e=fit$minimum
  }
  
  tree$edge.length <- tree$edge.length * sig2e
  ii <- sapply(1:Ntip(tree), function(x, e) which(e == x), 
               e = tree$edge[, 2])
  tree$edge.length[ii] <- tree$edge.length[ii] + se[tree$tip.label]^2
  vf <- diag(vcv(tree))[spp]
  w <- varFixed(~vf)
  obj <- gls(model, data = cbind(data, vf), correlation = corfunc(corClassValue, 
                                                                  tree, form = ~spp, ...), weights = w, method = method)
  obj$call <- Call
  obj$sig2e <- sig2e
  obj
}

#Internal function
pglsSEyPagelToOptimizeLambda=function(lambda,model,data,tree,...) {
  -logLik(modPgls.SEy(model=model,data=data,tree=tree,corClassValue=lambda,corClass=corPagel,fixed=T,...)) #Returns -logLikelihood of the pgls.SEy model with lambda fixed to the value of the lambda argument. sig2e will be optimized within modPgls.SEy unless given as an argument here
}

#Function intended for users
pglsSEyPagel=function(model, data, tree, lambdaInterval=c(0,1),...){
  optimizedModel=optimize(pglsSEyPagelToOptimizeLambda,lambdaInterval,model=model,data=data,tree=tree,...) #Optimizes lambda in the lambdaInterval using the pglsSEyPagelToOptimizeLambda function
  return(modPgls.SEy(model=model,data=data,tree=tree,corClass=corPagel,fixed=T,corClassValue=optimizedModel$minimum,...)) #Returns the final model fit
}

```

```{r}

model_inv_ges <- as.data.frame(model_inv_ges)
rownames(model_inv_ges) <- model_inv_ges$spp
se_inv_ges <- rep(0.01, nrow(model_inv_ges))
names(se_inv_ges) <- rownames(model_inv_ges)

model_inter_ges <- as.data.frame(model_inter_ges)
rownames(model_inter_ges) <- model_inter_ges$spp
se_inter_ges <- rep(0.01, nrow(model_inter_ges))
names(se_inter_ges) <- rownames(model_inter_ges)

model_inv_litter <- as.data.frame(model_inv_litter)
rownames(model_inv_litter) <- model_inv_litter$spp
se_inv_litter <- rep(0.01, nrow(model_inv_litter))
names(se_inv_litter) <- rownames(model_inv_litter)

model_inter_litter <- as.data.frame(model_inter_litter)
rownames(model_inter_litter) <- model_inter_litter$spp
se_inter_litter <- rep(0.01, nrow(model_inter_litter))
names(se_inter_litter) <- rownames(model_inter_litter)

model_inv_interval <- as.data.frame(model_inv_interval)
rownames(model_inv_interval) <- model_inv_interval$spp
se_inv_interval <- rep(0.01, nrow(model_inv_interval))
names(se_inv_interval) <- rownames(model_inv_interval)

model_inter_interval <- as.data.frame(model_inter_interval)
rownames(model_inter_interval) <- model_inter_interval$spp
se_inter_interval <- rep(0.01, nrow(model_inter_interval))
names(se_inter_interval) <- rownames(model_inter_interval)

model_inv_litsize <- as.data.frame(model_inv_litsize)
rownames(model_inv_litsize) <- model_inv_litsize$spp
se_inv_litsize <- rep(0.01, nrow(model_inv_litsize))
names(se_inv_litsize) <- rownames(model_inv_litsize)

model_inter_litsize <- as.data.frame(model_inter_litsize)
rownames(model_inter_litsize) <- model_inter_litsize$spp
se_inter_litsize <- rep(0.01, nrow(model_inter_litsize))
names(se_inter_litsize) <- rownames(model_inter_litsize)

model_inv_litsize <- as.data.frame(model_inv_litsize)
rownames(model_inv_litsize) <- model_inv_litsize$spp
se_inv_litsize <- rep(0.01, nrow(model_inv_litsize))
names(se_inv_litsize) <- rownames(model_inv_litsize)

model_inter_neo <- as.data.frame(model_inter_neo)
rownames(model_inter_neo) <- model_inter_neo$spp
se_inter_neo <- rep(0.01, nrow(model_inter_neo))
names(se_inter_neo) <- rownames(model_inter_neo)

model_inv_neo <- as.data.frame(model_inv_neo)
rownames(model_inv_neo) <- model_inv_neo$spp
se_inv_neo <- rep(0.01, nrow(model_inv_neo))
names(se_inv_neo) <- rownames(model_inv_neo)



# Female body mass subsets: gestation length
model_inv_ges_female <- as.data.frame(model_inv_ges_female)
rownames(model_inv_ges_female) <- model_inv_ges_female$spp
se_inv_ges_female <- rep(0.01, nrow(model_inv_ges_female))
names(se_inv_ges_female) <- rownames(model_inv_ges_female)

model_inter_ges_female <- as.data.frame(model_inter_ges_female)
rownames(model_inter_ges_female) <- model_inter_ges_female$spp
se_inter_ges_female <- rep(0.01, nrow(model_inter_ges_female))
names(se_inter_ges_female) <- rownames(model_inter_ges_female)


# Female body mass subsets: interbirth interval
model_inv_interval_female <- as.data.frame(model_inv_interval_female)
rownames(model_inv_interval_female) <- model_inv_interval_female$spp
se_inv_interval_female <- rep(0.01, nrow(model_inv_interval_female))
names(se_inv_interval_female) <- rownames(model_inv_interval_female)

model_inter_interval_female <- as.data.frame(model_inter_interval_female)
rownames(model_inter_interval_female) <- model_inter_interval_female$spp
se_inter_interval_female <- rep(0.01, nrow(model_inter_interval_female))
names(se_inter_interval_female) <- rownames(model_inter_interval_female)


# Female body mass subsets: litter mass
model_inv_litter_female <- as.data.frame(model_inv_litter_female)
rownames(model_inv_litter_female) <- model_inv_litter_female$spp
se_inv_litter_female <- rep(0.01, nrow(model_inv_litter_female))
names(se_inv_litter_female) <- rownames(model_inv_litter_female)

model_inter_litter_female <- as.data.frame(model_inter_litter_female)
rownames(model_inter_litter_female) <- model_inter_litter_female$spp
se_inter_litter_female <- rep(0.01, nrow(model_inter_litter_female))
names(se_inter_litter_female) <- rownames(model_inter_litter_female)


# Female body mass subsets: litter size
model_inv_litsize_female <- as.data.frame(model_inv_litsize_female)
rownames(model_inv_litsize_female) <- model_inv_litsize_female$spp
se_inv_litsize_female <- rep(0.01, nrow(model_inv_litsize_female))
names(se_inv_litsize_female) <- rownames(model_inv_litsize_female)

model_inter_litsize_female <- as.data.frame(model_inter_litsize_female)
rownames(model_inter_litsize_female) <- model_inter_litsize_female$spp
se_inter_litsize_female <- rep(0.01, nrow(model_inter_litsize_female))
names(se_inter_litsize_female) <- rownames(model_inter_litsize_female)


# Female body mass subsets: neonate body size
model_inv_neo_female <- as.data.frame(model_inv_neo_female)
rownames(model_inv_neo_female) <- model_inv_neo_female$spp
se_inv_neo_female <- rep(0.01, nrow(model_inv_neo_female))
names(se_inv_neo_female) <- rownames(model_inv_neo_female)

model_inter_neo_female <- as.data.frame(model_inter_neo_female)
rownames(model_inter_neo_female) <- model_inter_neo_female$spp
se_inter_neo_female <- rep(0.01, nrow(model_inter_neo_female))
names(se_inter_neo_female) <- rownames(model_inter_neo_female)





```

```{r gestation_length ~ invasiveness}

##GESTATION LENGTH

# Default factor with original reference
model_inv_ges$invasiveness <- factor(model_inv_ges$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

# Log transform variables
model_inv_ges$gestation_length <- log(model_inv_ges$gestation_length)
model_inv_ges$longevity <- log(model_inv_ges$longevity)
model_inv_ges$body_mass_g <- log(model_inv_ges$body_mass_g)

### NO COVARIATES
# Model 1: reference = Epitheliochorial
invas_ges_none_pgls_epi <- pglsSEyPagel(
  gestation_length ~ invasiveness,
  data = model_inv_ges,
  tree = comp_phylo_ges$phy,
  se = se_inv_ges,
  method = "ML"
)
summary(invas_ges_none_pgls_epi)

model_inv_ges$invasiveness <- relevel(model_inv_ges$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_ges_none_pgls_endo <- pglsSEyPagel(
  gestation_length ~ invasiveness,
  data = model_inv_ges,
  tree = comp_phylo_ges$phy,
  se = se_inv_ges,
  method = "ML"
)
summary(invas_ges_none_pgls_endo)

# Reset
model_inv_ges$invasiveness <- factor(model_inv_ges$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### LONGEVITY ONLY
# Model 1: reference = Epitheliochorial
invas_ges_long_pgls_epi <- pglsSEyPagel(
  gestation_length ~ invasiveness + longevity,
  data = model_inv_ges,
  tree = comp_phylo_ges$phy,
  se = se_inv_ges,
  method = "ML"
)
summary(invas_ges_long_pgls_epi)

model_inv_ges$invasiveness <- relevel(model_inv_ges$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_ges_long_pgls_endo <- pglsSEyPagel(
  gestation_length ~ invasiveness + longevity,
  data = model_inv_ges,
  tree = comp_phylo_ges$phy,
  se = se_inv_ges,
  method = "ML"
)
summary(invas_ges_long_pgls_endo)

# Reset
model_inv_ges$invasiveness <- factor(model_inv_ges$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### BODY MASS ONLY
# Model 1: reference = Epitheliochorial
invas_ges_adult_pgls_epi <- pglsSEyPagel(
  gestation_length ~ invasiveness + body_mass_g,
  data = model_inv_ges,
  tree = comp_phylo_ges$phy,
  se = se_inv_ges,
  method = "ML"
)
summary(invas_ges_adult_pgls_epi)

model_inv_ges$invasiveness <- relevel(model_inv_ges$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_ges_adult_pgls_endo <- pglsSEyPagel(
  gestation_length ~ invasiveness + body_mass_g,
  data = model_inv_ges,
  tree = comp_phylo_ges$phy,
  se = se_inv_ges,
  method = "ML"
)
summary(invas_ges_adult_pgls_endo)

# Reset
model_inv_ges$invasiveness <- factor(model_inv_ges$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### BOTH COVARIATES
# Model 1: reference = Epitheliochorial
invas_ges_both_pgls_epi <- pglsSEyPagel(
  gestation_length ~ invasiveness + longevity + body_mass_g,
  data = model_inv_ges,
  tree = comp_phylo_ges$phy,
  se = se_inv_ges,
  method = "ML"
)
summary(invas_ges_both_pgls_epi)

model_inv_ges$invasiveness <- relevel(model_inv_ges$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_ges_both_pgls_endo <- pglsSEyPagel(
  gestation_length ~ invasiveness + longevity + body_mass_g,
  data = model_inv_ges,
  tree = comp_phylo_ges$phy,
  se = se_inv_ges,
  method = "ML"
)
summary(invas_ges_both_pgls_endo)

# Reset
model_inv_ges$invasiveness <- factor(model_inv_ges$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

```

```{r gestation_length ~ interdigitation}

##GESTATION LENGTH

# Default factor with original reference
model_inter_ges$interdigitation <- factor(model_inter_ges$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

# Log transform variables
model_inter_ges$gestation_length <- log(model_inter_ges$gestation_length)
model_inter_ges$longevity <- log(model_inter_ges$longevity)
model_inter_ges$body_mass_g <- log(model_inter_ges$body_mass_g)

### NO COVARIATES
# Model 1: reference = Villous
inter_ges_none_pgls_vil <- pglsSEyPagel(
  gestation_length ~ interdigitation,
  data = model_inter_ges,
  tree = comp_phylo_ges$phy,
  se = se_inter_ges,
  method = "ML"
)
summary(inter_ges_none_pgls_vil)

model_inter_ges$interdigitation <- relevel(model_inter_ges$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_ges_none_pgls_trab <- pglsSEyPagel(
  gestation_length ~ interdigitation,
  data = model_inter_ges,
  tree = comp_phylo_ges$phy,
  se = se_inter_ges,
  method = "ML"
)
summary(inter_ges_none_pgls_trab)

# Reset
model_inter_ges$interdigitation <- factor(model_inter_ges$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### LONGEVITY ONLY
# Model 1: reference = Villous
inter_ges_long_pgls_vil <- pglsSEyPagel(
  gestation_length ~ interdigitation + longevity,
  data = model_inter_ges,
  tree = comp_phylo_ges$phy,
  se = se_inter_ges,
  method = "ML"
)
summary(inter_ges_long_pgls_vil)

model_inter_ges$interdigitation <- relevel(model_inter_ges$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_ges_long_pgls_trab <- pglsSEyPagel(
  gestation_length ~ interdigitation + longevity,
  data = model_inter_ges,
  tree = comp_phylo_ges$phy,
  se = se_inter_ges,
  method = "ML"
)
summary(inter_ges_long_pgls_trab)

# Reset
model_inter_ges$interdigitation <- factor(model_inter_ges$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### BODY MASS ONLY
# Model 1: reference = Villous
inter_ges_adult_pgls_vil <- pglsSEyPagel(
  gestation_length ~ interdigitation + body_mass_g,
  data = model_inter_ges,
  tree = comp_phylo_ges$phy,
  se = se_inter_ges,
  method = "ML"
)
summary(inter_ges_adult_pgls_vil)

model_inter_ges$interdigitation <- relevel(model_inter_ges$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_ges_adult_pgls_trab <- pglsSEyPagel(
  gestation_length ~ interdigitation + body_mass_g,
  data = model_inter_ges,
  tree = comp_phylo_ges$phy,
  se = se_inter_ges,
  method = "ML"
)
summary(inter_ges_adult_pgls_trab)

# Reset
model_inter_ges$interdigitation <- factor(model_inter_ges$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### BOTH COVARIATES
# Model 1: reference = Villous
inter_ges_both_pgls_vil <- pglsSEyPagel(
  gestation_length ~ interdigitation + longevity + body_mass_g,
  data = model_inter_ges,
  tree = comp_phylo_ges$phy,
  se = se_inter_ges,
  method = "ML"
)
summary(inter_ges_both_pgls_vil)

model_inter_ges$interdigitation <- relevel(model_inter_ges$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_ges_both_pgls_trab <- pglsSEyPagel(
  gestation_length ~ interdigitation + longevity + body_mass_g,
  data = model_inter_ges,
  tree = comp_phylo_ges$phy,
  se = se_inter_ges,
  method = "ML"
)
summary(inter_ges_both_pgls_trab)

# Reset
model_inter_ges$interdigitation <- factor(model_inter_ges$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

```

```{r litter_size ~ invasiveness}

##LITTER SIZE

# Default factor with original reference
model_inv_litsize$invasiveness <- factor(model_inv_litsize$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

# Log transform variables
model_inv_litsize$litter_size <- log(model_inv_litsize$litter_size)
model_inv_litsize$longevity <- log(model_inv_litsize$longevity)
model_inv_litsize$body_mass_g <- log(model_inv_litsize$body_mass_g)

### NO COVARIATES
# Model 1: reference = Epitheliochorial
invas_litsize_none_pgls_epi <- pglsSEyPagel(
  litter_size ~ invasiveness,
  data = model_inv_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inv_litsize,
  method = "ML"
)
summary(invas_litsize_none_pgls_epi)

model_inv_litsize$invasiveness <- relevel(model_inv_litsize$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_litsize_none_pgls_endo <- pglsSEyPagel(
  litter_size ~ invasiveness,
  data = model_inv_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inv_litsize,
  method = "ML"
)
summary(invas_litsize_none_pgls_endo)

# Reset
model_inv_litsize$invasiveness <- factor(model_inv_litsize$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### LONGEVITY ONLY
# Model 1: reference = Epitheliochorial
invas_litsize_long_pgls_epi <- pglsSEyPagel(
  litter_size ~ invasiveness + longevity,
  data = model_inv_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inv_litsize,
  method = "ML"
)
summary(invas_litsize_long_pgls_epi)

model_inv_litsize$invasiveness <- relevel(model_inv_litsize$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_litsize_long_pgls_endo <- pglsSEyPagel(
  litter_size ~ invasiveness + longevity,
  data = model_inv_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inv_litsize,
  method = "ML"
)
summary(invas_litsize_long_pgls_endo)

# Reset
model_inv_litsize$invasiveness <- factor(model_inv_litsize$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### BODY MASS ONLY
# Model 1: reference = Epitheliochorial
invas_litsize_adult_pgls_epi <- pglsSEyPagel(
  litter_size ~ invasiveness + body_mass_g,
  data = model_inv_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inv_litsize,
  method = "ML"
)
summary(invas_litsize_adult_pgls_epi)

model_inv_litsize$invasiveness <- relevel(model_inv_litsize$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_litsize_adult_pgls_endo <- pglsSEyPagel(
  litter_size ~ invasiveness + body_mass_g,
  data = model_inv_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inv_litsize,
  method = "ML"
)
summary(invas_litsize_adult_pgls_endo)

# Reset
model_inv_litsize$invasiveness <- factor(model_inv_litsize$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### BOTH COVARIATES
# Model 1: reference = Epitheliochorial
invas_litsize_both_pgls_epi <- pglsSEyPagel(
  litter_size ~ invasiveness + longevity + body_mass_g,
  data = model_inv_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inv_litsize,
  method = "ML"
)
summary(invas_litsize_both_pgls_epi)

model_inv_litsize$invasiveness <- relevel(model_inv_litsize$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_litsize_both_pgls_endo <- pglsSEyPagel(
  litter_size ~ invasiveness + longevity + body_mass_g,
  data = model_inv_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inv_litsize,
  method = "ML"
)
summary(invas_litsize_both_pgls_endo)

# Reset
model_inv_litsize$invasiveness <- factor(model_inv_litsize$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

```

```{r litter_size ~ interdigitation}

##LITTER SIZE

# Default factor with original reference
model_inter_litsize$interdigitation <- factor(model_inter_litsize$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

# Log transform variables
model_inter_litsize$litter_size <- log(model_inter_litsize$litter_size)
model_inter_litsize$longevity <- log(model_inter_litsize$longevity)
model_inter_litsize$body_mass_g <- log(model_inter_litsize$body_mass_g)

### NO COVARIATES
# Model 1: reference = Villous
inter_litsize_none_pgls_vil <- pglsSEyPagel(
  litter_size ~ interdigitation,
  data = model_inter_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inter_litsize,
  method = "ML"
)
summary(inter_litsize_none_pgls_vil)

model_inter_litsize$interdigitation <- relevel(model_inter_litsize$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_litsize_none_pgls_trab <- pglsSEyPagel(
  litter_size ~ interdigitation,
  data = model_inter_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inter_litsize,
  method = "ML"
)
summary(inter_litsize_none_pgls_trab)

# Reset
model_inter_litsize$interdigitation <- factor(model_inter_litsize$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### LONGEVITY ONLY
# Model 1: reference = Villous
inter_litsize_long_pgls_vil <- pglsSEyPagel(
  litter_size ~ interdigitation + longevity,
  data = model_inter_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inter_litsize,
  method = "ML"
)
summary(inter_litsize_long_pgls_vil)

model_inter_litsize$interdigitation <- relevel(model_inter_litsize$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_litsize_long_pgls_trab <- pglsSEyPagel(
  litter_size ~ interdigitation + longevity,
  data = model_inter_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inter_litsize,
  method = "ML"
)
summary(inter_litsize_long_pgls_trab)

# Reset
model_inter_litsize$interdigitation <- factor(model_inter_litsize$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### BODY MASS ONLY
# Model 1: reference = Villous
inter_litsize_adult_pgls_vil <- pglsSEyPagel(
  litter_size ~ interdigitation + body_mass_g,
  data = model_inter_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inter_litsize,
  method = "ML"
)
summary(inter_litsize_adult_pgls_vil)

model_inter_litsize$interdigitation <- relevel(model_inter_litsize$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_litsize_adult_pgls_trab <- pglsSEyPagel(
  litter_size ~ interdigitation + body_mass_g,
  data = model_inter_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inter_litsize,
  method = "ML"
)
summary(inter_litsize_adult_pgls_trab)

# Reset
model_inter_litsize$interdigitation <- factor(model_inter_litsize$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### BOTH COVARIATES
# Model 1: reference = Villous
inter_litsize_both_pgls_vil <- pglsSEyPagel(
  litter_size ~ interdigitation + longevity + body_mass_g,
  data = model_inter_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inter_litsize,
  method = "ML"
)
summary(inter_litsize_both_pgls_vil)

model_inter_litsize$interdigitation <- relevel(model_inter_litsize$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_litsize_both_pgls_trab <- pglsSEyPagel(
  litter_size ~ interdigitation + longevity + body_mass_g,
  data = model_inter_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inter_litsize,
  method = "ML"
)
summary(inter_litsize_both_pgls_trab)

# Reset
model_inter_litsize$interdigitation <- factor(model_inter_litsize$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

```

```{r relative_litter_mass ~ invasiveness}

##LITTER MASS

# Default factor with original reference
model_inv_litter$invasiveness <- factor(model_inv_litter$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

# Log transform variables
model_inv_litter$relative_litter_mass <- log(model_inv_litter$relative_litter_mass)
model_inv_litter$longevity <- log(model_inv_litter$longevity)
model_inv_litter$body_mass_g <- log(model_inv_litter$body_mass_g)

### NO COVARIATES
# Model 1: reference = Epitheliochorial
invas_lit_none_pgls_epi <- pglsSEyPagel(
  relative_litter_mass ~ invasiveness,
  data = model_inv_litter,
  tree = comp_phylo_lit$phy,
  se = se_inv_litter,
  method = "ML"
)
summary(invas_lit_none_pgls_epi)

model_inv_litter$invasiveness <- relevel(model_inv_litter$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_lit_none_pgls_endo <- pglsSEyPagel(
  relative_litter_mass ~ invasiveness,
  data = model_inv_litter,
  tree = comp_phylo_lit$phy,
  se = se_inv_litter,
  method = "ML"
)
summary(invas_lit_none_pgls_endo)

# Reset
model_inv_litter$invasiveness <- factor(model_inv_litter$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### LONGEVITY ONLY
# Model 1: reference = Epitheliochorial
invas_lit_long_pgls_epi <- pglsSEyPagel(
  relative_litter_mass ~ invasiveness + longevity,
  data = model_inv_litter,
  tree = comp_phylo_lit$phy,
  se = se_inv_litter,
  method = "ML"
)
summary(invas_lit_long_pgls_epi)

model_inv_litter$invasiveness <- relevel(model_inv_litter$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_lit_long_pgls_endo <- pglsSEyPagel(
  relative_litter_mass ~ invasiveness + longevity,
  data = model_inv_litter,
  tree = comp_phylo_lit$phy,
  se = se_inv_litter,
  method = "ML"
)
summary(invas_lit_long_pgls_endo)

# Reset
model_inv_litter$invasiveness <- factor(model_inv_litter$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))



```

```{r relative_litter_mass ~ interdigitation}

##LITTER MASS

# Default factor with original reference
model_inter_litter$interdigitation <- factor(model_inter_litter$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

# Log transform variables
model_inter_litter$relative_litter_mass <- log(model_inter_litter$relative_litter_mass)
model_inter_litter$longevity <- log(model_inter_litter$longevity)
model_inter_litter$body_mass_g <- log(model_inter_litter$body_mass_g)

### NO COVARIATES
# Model 1: reference = Villous
inter_lit_none_pgls_vil <- pglsSEyPagel(
  relative_litter_mass ~ interdigitation,
  data = model_inter_litter,
  tree = comp_phylo_lit$phy,
  se = se_inter_litter,
  method = "ML"
)
summary(inter_lit_none_pgls_vil)

model_inter_litter$interdigitation <- relevel(model_inter_litter$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_lit_none_pgls_trab <- pglsSEyPagel(
  relative_litter_mass ~ interdigitation,
  data = model_inter_litter,
  tree = comp_phylo_lit$phy,
  se = se_inter_litter,
  method = "ML"
)
summary(inter_lit_none_pgls_trab)

# Reset
model_inter_litter$interdigitation <- factor(model_inter_litter$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### LONGEVITY ONLY
# Model 1: reference = Villous
inter_lit_long_pgls_vil <- pglsSEyPagel(
  relative_litter_mass ~ interdigitation + longevity,
  data = model_inter_litter,
  tree = comp_phylo_lit$phy,
  se = se_inter_litter,
  method = "ML"
)
summary(inter_lit_long_pgls_vil)

model_inter_litter$interdigitation <- relevel(model_inter_litter$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_lit_long_pgls_trab <- pglsSEyPagel(
  relative_litter_mass ~ interdigitation + longevity,
  data = model_inter_litter,
  tree = comp_phylo_lit$phy,
  se = se_inter_litter,
  method = "ML"
)
summary(inter_lit_long_pgls_trab)

# Reset
model_inter_litter$interdigitation <- factor(model_inter_litter$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))



```

```{r neonate_over_adult ~ invasiveness}

##NEONATE BODY WEIGHT / ADULT BODY WEIGHT

# Default factor with original reference
model_inv_neo$invasiveness <- factor(model_inv_neo$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

# Log transform variables
model_inv_neo$neonate_over_adult <- log(model_inv_neo$neonate_over_adult)
model_inv_neo$longevity <- log(model_inv_neo$longevity)
model_inv_neo$body_mass_g <- log(model_inv_neo$body_mass_g)

### NO COVARIATES
# Model 1: reference = Epitheliochorial
invas_neo_none_pgls_epi <- pglsSEyPagel(
  neonate_over_adult ~ invasiveness,
  data = model_inv_neo,
  tree = comp_phylo_neo$phy,
  se = se_inv_neo,
  method = "ML"
)
summary(invas_neo_none_pgls_epi)

model_inv_neo$invasiveness <- relevel(model_inv_neo$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_neo_none_pgls_endo <- pglsSEyPagel(
  neonate_over_adult ~ invasiveness,
  data = model_inv_neo,
  tree = comp_phylo_neo$phy,
  se = se_inv_neo,
  method = "ML"
)
summary(invas_neo_none_pgls_endo)

# Reset
model_inv_neo$invasiveness <- factor(model_inv_neo$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### LONGEVITY ONLY
# Model 1: reference = Epitheliochorial
invas_neo_long_pgls_epi <- pglsSEyPagel(
  neonate_over_adult ~ invasiveness + longevity,
  data = model_inv_neo,
  tree = comp_phylo_neo$phy,
  se = se_inv_neo,
  method = "ML"
)
summary(invas_neo_long_pgls_epi)

model_inv_neo$invasiveness <- relevel(model_inv_neo$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_neo_long_pgls_endo <- pglsSEyPagel(
  neonate_over_adult ~ invasiveness + longevity,
  data = model_inv_neo,
  tree = comp_phylo_neo$phy,
  se = se_inv_neo,
  method = "ML"
)
summary(invas_neo_long_pgls_endo)

# Reset
model_inv_neo$invasiveness <- factor(model_inv_neo$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))



```

```{r neonate_over_adult ~ interdigitation}

##NEONATE BODY WEIGHT / ADULT BODY WEIGHT

# Default factor with original reference
model_inter_neo$interdigitation <- factor(model_inter_neo$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

# Log transform variables
model_inter_neo$neonate_over_adult <- log(model_inter_neo$neonate_over_adult)
model_inter_neo$longevity <- log(model_inter_neo$longevity)
model_inter_neo$body_mass_g <- log(model_inter_neo$body_mass_g)

### NO COVARIATES
# Model 1: reference = Villous
inter_neo_none_pgls_vil <- pglsSEyPagel(
  neonate_over_adult ~ interdigitation,
  data = model_inter_neo,
  tree = comp_phylo_neo$phy,
  se = se_inter_neo,
  method = "ML"
)
summary(inter_neo_none_pgls_vil)

model_inter_neo$interdigitation <- relevel(model_inter_neo$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_neo_none_pgls_trab <- pglsSEyPagel(
  neonate_over_adult ~ interdigitation,
  data = model_inter_neo,
  tree = comp_phylo_neo$phy,
  se = se_inter_neo,
  method = "ML"
)
summary(inter_neo_none_pgls_trab)

# Reset
model_inter_neo$interdigitation <- factor(model_inter_neo$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### LONGEVITY ONLY
# Model 1: reference = Villous
inter_neo_long_pgls_vil <- pglsSEyPagel(
  neonate_over_adult ~ interdigitation + longevity,
  data = model_inter_neo,
  tree = comp_phylo_neo$phy,
  se = se_inter_neo,
  method = "ML"
)
summary(inter_neo_long_pgls_vil)

model_inter_neo$interdigitation <- relevel(model_inter_neo$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_neo_long_pgls_trab <- pglsSEyPagel(
  neonate_over_adult ~ interdigitation + longevity,
  data = model_inter_neo,
  tree = comp_phylo_neo$phy,
  se = se_inter_neo,
  method = "ML"
)
summary(inter_neo_long_pgls_trab)

# Reset
model_inter_neo$interdigitation <- factor(model_inter_neo$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))



```

```{r interbirth_interval ~ invasiveness}

##INTERBIRTH INTERVAL

# Default factor with original reference
model_inv_interval$invasiveness <- factor(model_inv_interval$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

# Log transform variables
model_inv_interval$interbirth_interval <- log(model_inv_interval$interbirth_interval)
model_inv_interval$longevity <- log(model_inv_interval$longevity)
model_inv_interval$body_mass_g <- log(model_inv_interval$body_mass_g)

### NO COVARIATES
# Model 1: reference = Epitheliochorial
invas_interval_none_pgls_epi <- pglsSEyPagel(
  interbirth_interval ~ invasiveness,
  data = model_inv_interval,
  tree = comp_phylo_interval$phy,
  se = se_inv_interval,
  method = "ML"
)
summary(invas_interval_none_pgls_epi)

model_inv_interval$invasiveness <- relevel(model_inv_interval$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_interval_none_pgls_endo <- pglsSEyPagel(
  interbirth_interval ~ invasiveness,
  data = model_inv_interval,
  tree = comp_phylo_interval$phy,
  se = se_inv_interval,
  method = "ML"
)
summary(invas_interval_none_pgls_endo)

# Reset
model_inv_interval$invasiveness <- factor(model_inv_interval$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### LONGEVITY ONLY
# Model 1: reference = Epitheliochorial
invas_interval_long_pgls_epi <- pglsSEyPagel(
  interbirth_interval ~ invasiveness + longevity,
  data = model_inv_interval,
  tree = comp_phylo_interval$phy,
  se = se_inv_interval,
  method = "ML"
)
summary(invas_interval_long_pgls_epi)

model_inv_interval$invasiveness <- relevel(model_inv_interval$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_interval_long_pgls_endo <- pglsSEyPagel(
  interbirth_interval ~ invasiveness + longevity,
  data = model_inv_interval,
  tree = comp_phylo_interval$phy,
  se = se_inv_interval,
  method = "ML"
)
summary(invas_interval_long_pgls_endo)

# Reset
model_inv_interval$invasiveness <- factor(model_inv_interval$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### BODY MASS ONLY
# Model 1: reference = Epitheliochorial
invas_interval_adult_pgls_epi <- pglsSEyPagel(
  interbirth_interval ~ invasiveness + body_mass_g,
  data = model_inv_interval,
  tree = comp_phylo_interval$phy,
  se = se_inv_interval,
  method = "ML"
)
summary(invas_interval_adult_pgls_epi)

model_inv_interval$invasiveness <- relevel(model_inv_interval$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_interval_adult_pgls_endo <- pglsSEyPagel(
  interbirth_interval ~ invasiveness + body_mass_g,
  data = model_inv_interval,
  tree = comp_phylo_interval$phy,
  se = se_inv_interval,
  method = "ML"
)
summary(invas_interval_adult_pgls_endo)

# Reset
model_inv_interval$invasiveness <- factor(model_inv_interval$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### BOTH COVARIATES
# Model 1: reference = Epitheliochorial
invas_interval_both_pgls_epi <- pglsSEyPagel(
  interbirth_interval ~ invasiveness + longevity + body_mass_g,
  data = model_inv_interval,
  tree = comp_phylo_interval$phy,
  se = se_inv_interval,
  method = "ML"
)
summary(invas_interval_both_pgls_epi)

model_inv_interval$invasiveness <- relevel(model_inv_interval$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_interval_both_pgls_endo <- pglsSEyPagel(
  interbirth_interval ~ invasiveness + longevity + body_mass_g,
  data = model_inv_interval,
  tree = comp_phylo_interval$phy,
  se = se_inv_interval,
  method = "ML"
)
summary(invas_interval_both_pgls_endo)

# Reset
model_inv_interval$invasiveness <- factor(model_inv_interval$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

```

```{r interbirth_interval ~ interdigitation}

##INTERBIRTH INTERVAL

# Default factor with original reference
model_inter_interval$interdigitation <- factor(model_inter_interval$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

# Log transform variables
model_inter_interval$interbirth_interval <- log(model_inter_interval$interbirth_interval)
model_inter_interval$longevity <- log(model_inter_interval$longevity)
model_inter_interval$body_mass_g <- log(model_inter_interval$body_mass_g)

### NO COVARIATES
# Model 1: reference = Villous
inter_interval_none_pgls_vil <- pglsSEyPagel(
  interbirth_interval ~ interdigitation,
  data = model_inter_interval,
  tree = comp_phylo_interval$phy,
  se = se_inter_interval,
  method = "ML"
)
summary(inter_interval_none_pgls_vil)

model_inter_interval$interdigitation <- relevel(model_inter_interval$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_interval_none_pgls_trab <- pglsSEyPagel(
  interbirth_interval ~ interdigitation,
  data = model_inter_interval,
  tree = comp_phylo_interval$phy,
  se = se_inter_interval,
  method = "ML"
)
summary(inter_interval_none_pgls_trab)

# Reset
model_inter_interval$interdigitation <- factor(model_inter_interval$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### LONGEVITY ONLY
# Model 1: reference = Villous
inter_interval_long_pgls_vil <- pglsSEyPagel(
  interbirth_interval ~ interdigitation + longevity,
  data = model_inter_interval,
  tree = comp_phylo_interval$phy,
  se = se_inter_interval,
  method = "ML"
)
summary(inter_interval_long_pgls_vil)

model_inter_interval$interdigitation <- relevel(model_inter_interval$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_interval_long_pgls_trab <- pglsSEyPagel(
  interbirth_interval ~ interdigitation + longevity,
  data = model_inter_interval,
  tree = comp_phylo_interval$phy,
  se = se_inter_interval,
  method = "ML"
)
summary(inter_interval_long_pgls_trab)

# Reset
model_inter_interval$interdigitation <- factor(model_inter_interval$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### BODY MASS ONLY
# Model 1: reference = Villous
inter_interval_adult_pgls_vil <- pglsSEyPagel(
  interbirth_interval ~ interdigitation + body_mass_g,
  data = model_inter_interval,
  tree = comp_phylo_interval$phy,
  se = se_inter_interval,
  method = "ML"
)
summary(inter_interval_adult_pgls_vil)

model_inter_interval$interdigitation <- relevel(model_inter_interval$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_interval_adult_pgls_trab <- pglsSEyPagel(
  interbirth_interval ~ interdigitation + body_mass_g,
  data = model_inter_interval,
  tree = comp_phylo_interval$phy,
  se = se_inter_interval,
  method = "ML"
)
summary(inter_interval_adult_pgls_trab)

# Reset
model_inter_interval$interdigitation <- factor(model_inter_interval$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### BOTH COVARIATES
# Model 1: reference = Villous
inter_interval_both_pgls_vil <- pglsSEyPagel(
  interbirth_interval ~ interdigitation + longevity + body_mass_g,
  data = model_inter_interval,
  tree = comp_phylo_interval$phy,
  se = se_inter_interval,
  method = "ML"
)
summary(inter_interval_both_pgls_vil)

model_inter_interval$interdigitation <- relevel(model_inter_interval$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_interval_both_pgls_trab <- pglsSEyPagel(
  interbirth_interval ~ interdigitation + longevity + body_mass_g,
  data = model_inter_interval,
  tree = comp_phylo_interval$phy,
  se = se_inter_interval,
  method = "ML"
)
summary(inter_interval_both_pgls_trab)

# Reset
model_inter_interval$interdigitation <- factor(model_inter_interval$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

```

```{r female gestation_length ~ invasiveness}

##FEMALE SUBSET — GESTATION LENGTH



model_inv_ges_female$invasiveness <- factor(model_inv_ges_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

model_inv_ges_female$gestation_length <- log(model_inv_ges_female$gestation_length)
model_inv_ges_female$longevity <- log(model_inv_ges_female$longevity)
model_inv_ges_female$female_body_mass_g <- log(model_inv_ges_female$female_body_mass_g)

### FEMALE BODY MASS ONLY
# Model 1: reference = Epitheliochorial
invas_ges_female_pgls_epi <- pglsSEyPagel(
  gestation_length ~ invasiveness + female_body_mass_g,
  data = model_inv_ges_female,
  tree = comp_phylo_ges_female$phy,
  se = se_inv_ges_female,
  method = "ML"
)
summary(invas_ges_female_pgls_epi)

model_inv_ges_female$invasiveness <- relevel(model_inv_ges_female$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_ges_female_pgls_endo <- pglsSEyPagel(
  gestation_length ~ invasiveness + female_body_mass_g,
  data = model_inv_ges_female,
  tree = comp_phylo_ges_female$phy,
  se = se_inv_ges_female,
  method = "ML"
)
summary(invas_ges_female_pgls_endo)

# Reset
model_inv_ges_female$invasiveness <- factor(model_inv_ges_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### LONGEVITY + FEMALE BODY MASS
# Model 1: reference = Epitheliochorial
invas_ges_longfemale_pgls_epi <- pglsSEyPagel(
  gestation_length ~ invasiveness + longevity + female_body_mass_g,
  data = model_inv_ges_female,
  tree = comp_phylo_ges_female$phy,
  se = se_inv_ges_female,
  method = "ML"
)
summary(invas_ges_longfemale_pgls_epi)

model_inv_ges_female$invasiveness <- relevel(model_inv_ges_female$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_ges_longfemale_pgls_endo <- pglsSEyPagel(
  gestation_length ~ invasiveness + longevity + female_body_mass_g,
  data = model_inv_ges_female,
  tree = comp_phylo_ges_female$phy,
  se = se_inv_ges_female,
  method = "ML"
)
summary(invas_ges_longfemale_pgls_endo)

# Reset
model_inv_ges_female$invasiveness <- factor(model_inv_ges_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

```

```{r female gestation_length ~ interdigitation}

##FEMALE SUBSET — GESTATION LENGTH


model_inter_ges_female$interdigitation <- factor(model_inter_ges_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

model_inter_ges_female$gestation_length <- log(model_inter_ges_female$gestation_length)
model_inter_ges_female$longevity <- log(model_inter_ges_female$longevity)
model_inter_ges_female$female_body_mass_g <- log(model_inter_ges_female$female_body_mass_g)

### FEMALE BODY MASS ONLY
# Model 1: reference = Villous
inter_ges_female_pgls_vil <- pglsSEyPagel(
  gestation_length ~ interdigitation + female_body_mass_g,
  data = model_inter_ges_female,
  tree = comp_phylo_ges_female$phy,
  se = se_inter_ges_female,
  method = "ML"
)
summary(inter_ges_female_pgls_vil)

model_inter_ges_female$interdigitation <- relevel(model_inter_ges_female$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_ges_female_pgls_trab <- pglsSEyPagel(
  gestation_length ~ interdigitation + female_body_mass_g,
  data = model_inter_ges_female,
  tree = comp_phylo_ges_female$phy,
  se = se_inter_ges_female,
  method = "ML"
)
summary(inter_ges_female_pgls_trab)

# Reset
model_inter_ges_female$interdigitation <- factor(model_inter_ges_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### LONGEVITY + FEMALE BODY MASS
# Model 1: reference = Villous
inter_ges_longfemale_pgls_vil <- pglsSEyPagel(
  gestation_length ~ interdigitation + longevity + female_body_mass_g,
  data = model_inter_ges_female,
  tree = comp_phylo_ges_female$phy,
  se = se_inter_ges_female,
  method = "ML"
)
summary(inter_ges_longfemale_pgls_vil)

model_inter_ges_female$interdigitation <- relevel(model_inter_ges_female$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_ges_longfemale_pgls_trab <- pglsSEyPagel(
  gestation_length ~ interdigitation + longevity + female_body_mass_g,
  data = model_inter_ges_female,
  tree = comp_phylo_ges_female$phy,
  se = se_inter_ges_female,
  method = "ML"
)
summary(inter_ges_longfemale_pgls_trab)

# Reset
model_inter_ges_female$interdigitation <- factor(model_inter_ges_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

```

```{r female litter_size ~ invasiveness}

##FEMALE SUBSET — LITTER SIZE


model_inv_litsize_female$invasiveness <- factor(model_inv_litsize_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

model_inv_litsize_female$litter_size <- log(model_inv_litsize_female$litter_size)
model_inv_litsize_female$longevity <- log(model_inv_litsize_female$longevity)
model_inv_litsize_female$female_body_mass_g <- log(model_inv_litsize_female$female_body_mass_g)

### FEMALE BODY MASS ONLY
# Model 1: reference = Epitheliochorial
invas_litsize_female_pgls_epi <- pglsSEyPagel(
  litter_size ~ invasiveness + female_body_mass_g,
  data = model_inv_litsize_female,
  tree = comp_phylo_litsize_female$phy,
  se = se_inv_litsize_female,
  method = "ML"
)
summary(invas_litsize_female_pgls_epi)

model_inv_litsize_female$invasiveness <- relevel(model_inv_litsize_female$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_litsize_female_pgls_endo <- pglsSEyPagel(
  litter_size ~ invasiveness + female_body_mass_g,
  data = model_inv_litsize_female,
  tree = comp_phylo_litsize_female$phy,
  se = se_inv_litsize_female,
  method = "ML"
)
summary(invas_litsize_female_pgls_endo)

# Reset
model_inv_litsize_female$invasiveness <- factor(model_inv_litsize_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### LONGEVITY + FEMALE BODY MASS
# Model 1: reference = Epitheliochorial
invas_litsize_longfemale_pgls_epi <- pglsSEyPagel(
  litter_size ~ invasiveness + longevity + female_body_mass_g,
  data = model_inv_litsize_female,
  tree = comp_phylo_litsize_female$phy,
  se = se_inv_litsize_female,
  method = "ML"
)
summary(invas_litsize_longfemale_pgls_epi)

model_inv_litsize_female$invasiveness <- relevel(model_inv_litsize_female$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_litsize_longfemale_pgls_endo <- pglsSEyPagel(
  litter_size ~ invasiveness + longevity + female_body_mass_g,
  data = model_inv_litsize_female,
  tree = comp_phylo_litsize_female$phy,
  se = se_inv_litsize_female,
  method = "ML"
)
summary(invas_litsize_longfemale_pgls_endo)

# Reset
model_inv_litsize_female$invasiveness <- factor(model_inv_litsize_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

```

```{r female litter_size ~ interdigitation}

##FEMALE SUBSET — LITTER SIZE


model_inter_litsize_female$interdigitation <- factor(model_inter_litsize_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

model_inter_litsize_female$litter_size <- log(model_inter_litsize_female$litter_size)
model_inter_litsize_female$longevity <- log(model_inter_litsize_female$longevity)
model_inter_litsize_female$female_body_mass_g <- log(model_inter_litsize_female$female_body_mass_g)

### FEMALE BODY MASS ONLY
# Model 1: reference = Villous
inter_litsize_female_pgls_vil <- pglsSEyPagel(
  litter_size ~ interdigitation + female_body_mass_g,
  data = model_inter_litsize_female,
  tree = comp_phylo_litsize_female$phy,
  se = se_inter_litsize_female,
  method = "ML"
)
summary(inter_litsize_female_pgls_vil)

model_inter_litsize_female$interdigitation <- relevel(model_inter_litsize_female$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_litsize_female_pgls_trab <- pglsSEyPagel(
  litter_size ~ interdigitation + female_body_mass_g,
  data = model_inter_litsize_female,
  tree = comp_phylo_litsize_female$phy,
  se = se_inter_litsize_female,
  method = "ML"
)
summary(inter_litsize_female_pgls_trab)

# Reset
model_inter_litsize_female$interdigitation <- factor(model_inter_litsize_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### LONGEVITY + FEMALE BODY MASS
# Model 1: reference = Villous
inter_litsize_longfemale_pgls_vil <- pglsSEyPagel(
  litter_size ~ interdigitation + longevity + female_body_mass_g,
  data = model_inter_litsize_female,
  tree = comp_phylo_litsize_female$phy,
  se = se_inter_litsize_female,
  method = "ML"
)
summary(inter_litsize_longfemale_pgls_vil)

model_inter_litsize_female$interdigitation <- relevel(model_inter_litsize_female$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_litsize_longfemale_pgls_trab <- pglsSEyPagel(
  litter_size ~ interdigitation + longevity + female_body_mass_g,
  data = model_inter_litsize_female,
  tree = comp_phylo_litsize_female$phy,
  se = se_inter_litsize_female,
  method = "ML"
)
summary(inter_litsize_longfemale_pgls_trab)

# Reset
model_inter_litsize_female$interdigitation <- factor(model_inter_litsize_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

```

```{r female relative_litter_mass ~ invasiveness}

##FEMALE SUBSET — LITTER MASS

model_inv_litter_female$invasiveness <- factor(model_inv_litter_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

model_inv_litter_female$relative_litter_mass_female <- log(model_inv_litter_female$relative_litter_mass_female)
model_inv_litter_female$longevity <- log(model_inv_litter_female$longevity)
model_inv_litter_female$female_body_mass_g <- log(model_inv_litter_female$female_body_mass_g)

### FEMALE BODY MASS ONLY
# Model 1: reference = Epitheliochorial
invas_lit_female_pgls_epi <- pglsSEyPagel(
  relative_litter_mass_female ~ invasiveness,
  data = model_inv_litter_female,
  tree = comp_phylo_lit_female$phy,
  se = se_inv_litter_female,
  method = "ML"
)
summary(invas_lit_female_pgls_epi)

model_inv_litter_female$invasiveness <- relevel(model_inv_litter_female$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_lit_female_pgls_endo <- pglsSEyPagel(
  relative_litter_mass_female ~ invasiveness,
  data = model_inv_litter_female,
  tree = comp_phylo_lit_female$phy,
  se = se_inv_litter_female,
  method = "ML"
)
summary(invas_lit_female_pgls_endo)

# Reset
model_inv_litter_female$invasiveness <- factor(model_inv_litter_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### LONGEVITY
# Model 1: reference = Epitheliochorial
invas_lit_longfemale_pgls_epi <- pglsSEyPagel(
  relative_litter_mass_female ~ invasiveness + longevity,
  data = model_inv_litter_female,
  tree = comp_phylo_lit_female$phy,
  se = se_inv_litter_female,
  method = "ML"
)
summary(invas_lit_longfemale_pgls_epi)

model_inv_litter_female$invasiveness <- relevel(model_inv_litter_female$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_lit_longfemale_pgls_endo <- pglsSEyPagel(
  relative_litter_mass_female ~ invasiveness + longevity,
  data = model_inv_litter_female,
  tree = comp_phylo_lit_female$phy,
  se = se_inv_litter_female,
  method = "ML"
)
summary(invas_lit_longfemale_pgls_endo)

# Reset
model_inv_litter_female$invasiveness <- factor(model_inv_litter_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

```

```{r female relative_litter_mass ~ interdigitation}

##FEMALE SUBSET — LITTER MASS

model_inter_litter_female$interdigitation <- factor(model_inter_litter_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

model_inter_litter_female$relative_litter_mass_female <- log(model_inter_litter_female$relative_litter_mass_female)
model_inter_litter_female$longevity <- log(model_inter_litter_female$longevity)
model_inter_litter_female$female_body_mass_g <- log(model_inter_litter_female$female_body_mass_g)

### FEMALE BODY MASS ONLY
# Model 1: reference = Villous
inter_lit_female_pgls_vil <- pglsSEyPagel(
  relative_litter_mass_female ~ interdigitation,
  data = model_inter_litter_female,
  tree = comp_phylo_lit_female$phy,
  se = se_inter_litter_female,
  method = "ML"
)
summary(inter_lit_female_pgls_vil)

model_inter_litter_female$interdigitation <- relevel(model_inter_litter_female$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_lit_female_pgls_trab <- pglsSEyPagel(
  relative_litter_mass_female ~ interdigitation,
  data = model_inter_litter_female,
  tree = comp_phylo_lit_female$phy,
  se = se_inter_litter_female,
  method = "ML"
)
summary(inter_lit_female_pgls_trab)

# Reset
model_inter_litter_female$interdigitation <- factor(model_inter_litter_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### LONGEVITY 
# Model 1: reference = Villous
inter_lit_longfemale_pgls_vil <- pglsSEyPagel(
  relative_litter_mass_female ~ interdigitation + longevity,
  data = model_inter_litter_female,
  tree = comp_phylo_lit_female$phy,
  se = se_inter_litter_female,
  method = "ML"
)
summary(inter_lit_longfemale_pgls_vil)

model_inter_litter_female$interdigitation <- relevel(model_inter_litter_female$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_lit_longfemale_pgls_trab <- pglsSEyPagel(
  relative_litter_mass_female ~ interdigitation + longevity,
  data = model_inter_litter_female,
  tree = comp_phylo_lit_female$phy,
  se = se_inter_litter_female,
  method = "ML"
)
summary(inter_lit_longfemale_pgls_trab)

# Reset
model_inter_litter_female$interdigitation <- factor(model_inter_litter_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

```

```{r female neonate_over_adult ~ invasiveness}

##FEMALE SUBSET — NEONATE BODY WEIGHT / FEMALE BODY WEIGHT

model_inv_neo_female$invasiveness <- factor(model_inv_neo_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

model_inv_neo_female$neonate_over_adult <- log(model_inv_neo_female$neonate_over_adult)
model_inv_neo_female$longevity <- log(model_inv_neo_female$longevity)
model_inv_neo_female$female_body_mass_g <- log(model_inv_neo_female$female_body_mass_g)

### FEMALE BODY MASS ONLY
# Model 1: reference = Epitheliochorial
invas_neo_female_pgls_epi <- pglsSEyPagel(
  neonate_over_adult ~ invasiveness,
  data = model_inv_neo_female,
  tree = comp_phylo_neo_female$phy,
  se = se_inv_neo_female,
  method = "ML"
)
summary(invas_neo_female_pgls_epi)

model_inv_neo_female$invasiveness <- relevel(model_inv_neo_female$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_neo_female_pgls_endo <- pglsSEyPagel(
  neonate_over_adult ~ invasiveness,
  data = model_inv_neo_female,
  tree = comp_phylo_neo_female$phy,
  se = se_inv_neo_female,
  method = "ML"
)
summary(invas_neo_female_pgls_endo)

# Reset
model_inv_neo_female$invasiveness <- factor(model_inv_neo_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### LONGEVITY + FEMALE BODY MASS
# Model 1: reference = Epitheliochorial
invas_neo_longfemale_pgls_epi <- pglsSEyPagel(
  neonate_over_adult ~ invasiveness + longevity,
  data = model_inv_neo_female,
  tree = comp_phylo_neo_female$phy,
  se = se_inv_neo_female,
  method = "ML"
)
summary(invas_neo_longfemale_pgls_epi)

model_inv_neo_female$invasiveness <- relevel(model_inv_neo_female$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_neo_longfemale_pgls_endo <- pglsSEyPagel(
  neonate_over_adult ~ invasiveness + longevity,
  data = model_inv_neo_female,
  tree = comp_phylo_neo_female$phy,
  se = se_inv_neo_female,
  method = "ML"
)
summary(invas_neo_longfemale_pgls_endo)

# Reset
model_inv_neo_female$invasiveness <- factor(model_inv_neo_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

```

```{r female neonate_over_adult ~ interdigitation}

##FEMALE SUBSET — NEONATE BODY WEIGHT / FEMALE BODY WEIGHT

model_inter_neo_female$interdigitation <- factor(model_inter_neo_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

model_inter_neo_female$neonate_over_adult <- log(model_inter_neo_female$neonate_over_adult)
model_inter_neo_female$longevity <- log(model_inter_neo_female$longevity)
model_inter_neo_female$female_body_mass_g <- log(model_inter_neo_female$female_body_mass_g)

### FEMALE BODY MASS ONLY
# Model 1: reference = Villous
inter_neo_female_pgls_vil <- pglsSEyPagel(
  neonate_over_adult ~ interdigitation,
  data = model_inter_neo_female,
  tree = comp_phylo_neo_female$phy,
  se = se_inter_neo_female,
  method = "ML"
)
summary(inter_neo_female_pgls_vil)

model_inter_neo_female$interdigitation <- relevel(model_inter_neo_female$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_neo_female_pgls_trab <- pglsSEyPagel(
  neonate_over_adult ~ interdigitation,
  data = model_inter_neo_female,
  tree = comp_phylo_neo_female$phy,
  se = se_inter_neo_female,
  method = "ML"
)
summary(inter_neo_female_pgls_trab)

# Reset
model_inter_neo_female$interdigitation <- factor(model_inter_neo_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### LONGEVITY + FEMALE BODY MASS
# Model 1: reference = Villous
inter_neo_longfemale_pgls_vil <- pglsSEyPagel(
  neonate_over_adult ~ interdigitation + longevity,
  data = model_inter_neo_female,
  tree = comp_phylo_neo_female$phy,
  se = se_inter_neo_female,
  method = "ML"
)
summary(inter_neo_longfemale_pgls_vil)

model_inter_neo_female$interdigitation <- relevel(model_inter_neo_female$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_neo_longfemale_pgls_trab <- pglsSEyPagel(
  neonate_over_adult ~ interdigitation + longevity,
  data = model_inter_neo_female,
  tree = comp_phylo_neo_female$phy,
  se = se_inter_neo_female,
  method = "ML"
)
summary(inter_neo_longfemale_pgls_trab)

# Reset
model_inter_neo_female$interdigitation <- factor(model_inter_neo_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

```

```{r female interbirth_interval ~ invasiveness}

##FEMALE SUBSET — INTERBIRTH INTERVAL

model_inv_interval_female$invasiveness <- factor(model_inv_interval_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

model_inv_interval_female$interbirth_interval <- log(model_inv_interval_female$interbirth_interval)
model_inv_interval_female$longevity <- log(model_inv_interval_female$longevity)
model_inv_interval_female$female_body_mass_g <- log(model_inv_interval_female$female_body_mass_g)

### FEMALE BODY MASS ONLY
# Model 1: reference = Epitheliochorial
invas_interval_female_pgls_epi <- pglsSEyPagel(
  interbirth_interval ~ invasiveness + female_body_mass_g,
  data = model_inv_interval_female,
  tree = comp_phylo_interval_female$phy,
  se = se_inv_interval_female,
  method = "ML"
)
summary(invas_interval_female_pgls_epi)

model_inv_interval_female$invasiveness <- relevel(model_inv_interval_female$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_interval_female_pgls_endo <- pglsSEyPagel(
  interbirth_interval ~ invasiveness + female_body_mass_g,
  data = model_inv_interval_female,
  tree = comp_phylo_interval_female$phy,
  se = se_inv_interval_female,
  method = "ML"
)
summary(invas_interval_female_pgls_endo)

# Reset
model_inv_interval_female$invasiveness <- factor(model_inv_interval_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

### LONGEVITY + FEMALE BODY MASS
# Model 1: reference = Epitheliochorial
invas_interval_longfemale_pgls_epi <- pglsSEyPagel(
  interbirth_interval ~ invasiveness + longevity + female_body_mass_g,
  data = model_inv_interval_female,
  tree = comp_phylo_interval_female$phy,
  se = se_inv_interval_female,
  method = "ML"
)
summary(invas_interval_longfemale_pgls_epi)

model_inv_interval_female$invasiveness <- relevel(model_inv_interval_female$invasiveness, ref = "Endotheliochorial")

# Model 2: reference = Endotheliochorial
invas_interval_longfemale_pgls_endo <- pglsSEyPagel(
  interbirth_interval ~ invasiveness + longevity + female_body_mass_g,
  data = model_inv_interval_female,
  tree = comp_phylo_interval_female$phy,
  se = se_inv_interval_female,
  method = "ML"
)
summary(invas_interval_longfemale_pgls_endo)

# Reset
model_inv_interval_female$invasiveness <- factor(model_inv_interval_female$invasiveness, levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

```

```{r female interbirth_interval ~ interdigitation}

##FEMALE SUBSET — INTERBIRTH INTERVAL

model_inter_interval_female$interdigitation <- factor(model_inter_interval_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

model_inter_interval_female$interbirth_interval <- log(model_inter_interval_female$interbirth_interval)
model_inter_interval_female$longevity <- log(model_inter_interval_female$longevity)
model_inter_interval_female$female_body_mass_g <- log(model_inter_interval_female$female_body_mass_g)

### FEMALE BODY MASS ONLY
# Model 1: reference = Villous
inter_interval_female_pgls_vil <- pglsSEyPagel(
  interbirth_interval ~ interdigitation + female_body_mass_g,
  data = model_inter_interval_female,
  tree = comp_phylo_interval_female$phy,
  se = se_inter_interval_female,
  method = "ML"
)
summary(inter_interval_female_pgls_vil)

model_inter_interval_female$interdigitation <- relevel(model_inter_interval_female$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_interval_female_pgls_trab <- pglsSEyPagel(
  interbirth_interval ~ interdigitation + female_body_mass_g,
  data = model_inter_interval_female,
  tree = comp_phylo_interval_female$phy,
  se = se_inter_interval_female,
  method = "ML"
)
summary(inter_interval_female_pgls_trab)

# Reset
model_inter_interval_female$interdigitation <- factor(model_inter_interval_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

### LONGEVITY + FEMALE BODY MASS
# Model 1: reference = Villous
inter_interval_longfemale_pgls_vil <- pglsSEyPagel(
  interbirth_interval ~ interdigitation + longevity + female_body_mass_g,
  data = model_inter_interval_female,
  tree = comp_phylo_interval_female$phy,
  se = se_inter_interval_female,
  method = "ML"
)
summary(inter_interval_longfemale_pgls_vil)

model_inter_interval_female$interdigitation <- relevel(model_inter_interval_female$interdigitation, ref = "Trabecular")

# Model 2: reference = Trabecular
inter_interval_longfemale_pgls_trab <- pglsSEyPagel(
  interbirth_interval ~ interdigitation + longevity + female_body_mass_g,
  data = model_inter_interval_female,
  tree = comp_phylo_interval_female$phy,
  se = se_inter_interval_female,
  method = "ML"
)
summary(inter_interval_longfemale_pgls_trab)

# Reset
model_inter_interval_female$interdigitation <- factor(model_inter_interval_female$interdigitation, levels = c("Villous", "Trabecular", "Labyrinthine"))

```
```{r}
# ============================================================
# BASELINE PGLS MODELS WITHOUT PLACENTATION
# Allometric models for reviewer comparison
# ============================================================


# ------------------------------------------------------------
# Gestation length
# ------------------------------------------------------------

ges_none_pgls <- pglsSEyPagel(
  gestation_length ~ 1,
  data = model_inv_ges,
  tree = comp_phylo_ges$phy,
  se = se_inv_ges,
  method = "ML"
)

ges_adult_pgls <- pglsSEyPagel(
  gestation_length ~ body_mass_g,
  data = model_inv_ges,
  tree = comp_phylo_ges$phy,
  se = se_inv_ges,
  method = "ML"
)

ges_long_pgls <- pglsSEyPagel(
  gestation_length ~ longevity,
  data = model_inv_ges,
  tree = comp_phylo_ges$phy,
  se = se_inv_ges,
  method = "ML"
)

ges_long_adult_pgls <- pglsSEyPagel(
  gestation_length ~ longevity + body_mass_g,
  data = model_inv_ges,
  tree = comp_phylo_ges$phy,
  se = se_inv_ges,
  method = "ML"
)

ges_female_pgls <- pglsSEyPagel(
  gestation_length ~ female_body_mass_g,
  data = model_inv_ges_female,
  tree = comp_phylo_ges_female$phy,
  se = se_inv_ges_female,
  method = "ML"
)


# ------------------------------------------------------------
# Interbirth interval
# ------------------------------------------------------------

ibi_none_pgls <- pglsSEyPagel(
  interbirth_interval ~ 1,
  data = model_inv_interval,
  tree = comp_phylo_interval$phy,
  se = se_inv_interval,
  method = "ML"
)

ibi_adult_pgls <- pglsSEyPagel(
  interbirth_interval ~ body_mass_g,
  data = model_inv_interval,
  tree = comp_phylo_interval$phy,
  se = se_inv_interval,
  method = "ML"
)

ibi_long_pgls <- pglsSEyPagel(
  interbirth_interval ~ longevity,
  data = model_inv_interval,
  tree = comp_phylo_interval$phy,
  se = se_inv_interval,
  method = "ML"
)

ibi_long_adult_pgls <- pglsSEyPagel(
  interbirth_interval ~ longevity + body_mass_g,
  data = model_inv_interval,
  tree = comp_phylo_interval$phy,
  se = se_inv_interval,
  method = "ML"
)

ibi_female_pgls <- pglsSEyPagel(
  interbirth_interval ~ female_body_mass_g,
  data = model_inv_interval_female,
  tree = comp_phylo_interval_female$phy,
  se = se_inv_interval_female,
  method = "ML"
)

# ------------------------------------------------------------
# Litter mass
# ------------------------------------------------------------

litmass_none_pgls <- pglsSEyPagel(
  relative_litter_mass ~ 1,
  data = model_inv_litter,
  tree = comp_phylo_lit$phy,
  se = se_inv_litter,
  method = "ML"
)

litmass_long_pgls <- pglsSEyPagel(
  relative_litter_mass ~ longevity,
  data = model_inv_litter,
  tree = comp_phylo_lit$phy,
  se = se_inv_litter,
  method = "ML"
)

litmass_female_pgls <- pglsSEyPagel(
  relative_litter_mass_female ~ 1,
  data = model_inv_litter_female,
  tree = comp_phylo_lit_female$phy,
  se = se_inv_litter_female,
  method = "ML"
)

# ------------------------------------------------------------
# Litter size#
# ------------------------------------------------------------


litsize_none_pgls <- pglsSEyPagel(
  litter_size ~ 1,
  data = model_inv_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inv_litsize,
  method = "ML"
)

litsize_adult_pgls <- pglsSEyPagel(
  litter_size ~ body_mass_g,
  data = model_inv_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inv_litsize,
  method = "ML"
)

litsize_long_pgls <- pglsSEyPagel(
  litter_size ~ longevity,
  data = model_inv_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inv_litsize,
  method = "ML"
)

litsize_long_adult_pgls <- pglsSEyPagel(
  litter_size ~ longevity + body_mass_g,
  data = model_inv_litsize,
  tree = comp_phylo_litsize$phy,
  se = se_inv_litsize,
  method = "ML"
)

litsize_female_pgls <- pglsSEyPagel(
  litter_size ~ female_body_mass_g,
  data = model_inv_litsize_female,
  tree = comp_phylo_litsize_female$phy,
  se = se_inv_litsize_female,
  method = "ML"
)

# ------------------------------------------------------------
# Neonate body size

# ------------------------------------------------------------

neo_none_pgls <- pglsSEyPagel(
  neonate_over_adult ~ 1,
  data = model_inv_neo,
  tree = comp_phylo_neo$phy,
  se = se_inv_neo,
  method = "ML"
)


neo_long_pgls <- pglsSEyPagel(
  neonate_over_adult ~ longevity,
  data = model_inv_neo,
  tree = comp_phylo_neo$phy,
  se = se_inv_neo,
  method = "ML"
)

neo_female_pgls <- pglsSEyPagel(
  neonate_over_adult ~ 1,
  data = model_inv_neo_female,
  tree = comp_phylo_neo_female$phy,
  se = se_inv_neo_female,
  method = "ML"
)

#models previously missing

ges_longfemale_pgls <- pglsSEyPagel(
  gestation_length ~ longevity + female_body_mass_g,
  data = model_inv_ges_female,
  tree = comp_phylo_ges_female$phy,
  se = se_inv_ges_female,
  method = "ML"
)

ibi_longfemale_pgls <- pglsSEyPagel(
  interbirth_interval ~ longevity + female_body_mass_g,
  data = model_inv_interval_female,
  tree = comp_phylo_interval_female$phy,
  se = se_inv_interval_female,
  method = "ML"
)


litsize_longfemale_pgls <- pglsSEyPagel(
  litter_size ~ longevity + female_body_mass_g,
  data = model_inv_litsize_female,
  tree = comp_phylo_litsize_female$phy,
  se = se_inv_litsize_female,
  method = "ML"
)

```

```{r R2 extraction}
# ============================================================
# 1. Collect model names
# ============================================================

model_names_main <- ls(
  pattern = "^(invas|inter)_(ges|litsize|lit|neo|interval)_(none|long|adult|both|female|longfemale)_pgls_(epi|endo|hemo|vil|trab|laby)$"
)

model_names_base <- c(
  "ges_none_pgls",
  "ges_adult_pgls",
  "ges_long_pgls",
  "ges_long_adult_pgls",
  "ges_female_pgls",
  "ges_longfemale_pgls",

  "ibi_none_pgls",
  "ibi_adult_pgls",
  "ibi_long_pgls",
  "ibi_long_adult_pgls",
  "ibi_female_pgls",
  "ibi_longfemale_pgls",

  "litmass_none_pgls",
  "litmass_long_pgls",
  "litmass_adult_pgls",
  "litmass_long_adult_pgls",
  "litmass_female_pgls",
  "litmass_longfemale_pgls",

  "litsize_none_pgls",
  "litsize_adult_pgls",
  "litsize_long_pgls",
  "litsize_long_adult_pgls",
  "litsize_female_pgls",
  "litsize_longfemale_pgls",

  "neo_none_pgls",
  "neo_long_pgls",
  "neo_adult_pgls",
  "neo_long_adult_pgls",
  "neo_female_pgls",
  "neo_longfemale_pgls"
)

model_names_base <- model_names_base[model_names_base %in% ls()]

model_names <- c(model_names_main, model_names_base)

# ============================================================
# 2. Extract lambda
# ============================================================

extract_lambda <- function(model) {
  model_summary <- summary(model)

  if (!is.null(model_summary$modelStruct$corStruct)) {
    return(as.numeric(model_summary$modelStruct$corStruct))
  }

  NA_real_
}

# ============================================================
# 3. Match each full model to its baseline model
# ============================================================

get_baseline_model <- function(trait, covariates) {

  if (trait == "Gestation length") {
    if (covariates == "None") return("ges_none_pgls")
    if (covariates == "Adult body mass") return("ges_adult_pgls")
    if (covariates == "Longevity + adult body mass") return("ges_long_adult_pgls")
    if (covariates == "Female body mass") return("ges_female_pgls")
    if (covariates == "Longevity + female body mass") return("ges_longfemale_pgls")
  }

  if (trait == "Interbirth interval") {
    if (covariates == "None") return("ibi_none_pgls")
    if (covariates == "Adult body mass") return("ibi_adult_pgls")
    if (covariates == "Longevity + adult body mass") return("ibi_long_adult_pgls")
    if (covariates == "Female body mass") return("ibi_female_pgls")
    if (covariates == "Longevity + female body mass") return("ibi_longfemale_pgls")
  }

  if (trait == "Relative litter mass") {
    if (covariates == "None") return("litmass_none_pgls")
    if (covariates == "Longevity") return("litmass_long_pgls")
    if (covariates == "Adult body mass") return("litmass_adult_pgls")
    if (covariates == "Longevity + adult body mass") return("litmass_long_adult_pgls")
    if (covariates == "Female body mass") return("litmass_female_pgls")
    if (covariates == "Longevity + female body mass") return("litmass_longfemale_pgls")
  }

  if (trait == "Litter size") {
    if (covariates == "None") return("litsize_none_pgls")
    if (covariates == "Adult body mass") return("litsize_adult_pgls")
    if (covariates == "Longevity + adult body mass") return("litsize_long_adult_pgls")
    if (covariates == "Female body mass") return("litsize_female_pgls")
    if (covariates == "Longevity + female body mass") return("litsize_longfemale_pgls")
  }

  if (trait == "Neonate body weight / adult body weight") {
    if (covariates == "None") return("neo_none_pgls")
    if (covariates == "Longevity") return("neo_long_pgls")
    if (covariates == "Adult body mass") return("neo_adult_pgls")
    if (covariates == "Longevity + adult body mass") return("neo_long_adult_pgls")
    if (covariates == "Female body mass") return("neo_female_pgls")
    if (covariates == "Longevity + female body mass") return("neo_longfemale_pgls")
  }

  return(NA_character_)
}

# ============================================================
# 4. Compute rr2 R2_pred partial R²
# ============================================================

compute_partial_r2 <- function(model_name, trait, covariates, factor) {

  if (factor == "None (allometric model)") return(NA_real_)

  baseline_name <- get_baseline_model(trait, covariates)

  if (is.na(baseline_name)) return(NA_real_)
  if (!exists(baseline_name)) return(NA_real_)

  model_full <- get(model_name)
  model_base <- get(baseline_name)

  if (nobs(model_full) != nobs(model_base)) {
    warning(paste(
      "Different nobs for", model_name, "and", baseline_name,
      "- rr2 R2_pred set to NA"
    ))
    return(NA_real_)
  }

  as.numeric(rr2::R2_pred(model_full, model_base))
}

# ============================================================
# 5. Parse model names
# ============================================================

parse_model_name <- function(model_name) {

  if (model_name %in% model_names_base) {

    parts <- strsplit(model_name, "_")[[1]]

    trait <- dplyr::recode(
      parts[1],
      ges = "Gestation length",
      ibi = "Interbirth interval",
      litmass = "Relative litter mass",
      litsize = "Litter size",
      neo = "Neonate body weight / adult body weight"
    )

    covariate_code <- paste(parts[2:(length(parts) - 1)], collapse = "_")

    covariates <- dplyr::recode(
      covariate_code,
      none = "None",
      long = "Longevity",
      adult = "Adult body mass",
      long_adult = "Longevity + adult body mass",
      female = "Female body mass",
      longfemale = "Longevity + female body mass"
    )

    return(tibble::tibble(
      model_name = model_name,
      factor = "None (allometric model)",
      trait = trait,
      covariates = covariates,
      reference = NA_character_
    ))
  }

  parts <- strsplit(model_name, "_")[[1]]

  tibble::tibble(
    model_name = model_name,

    factor = dplyr::recode(
      parts[1],
      invas = "Invasiveness",
      inter = "Interdigitation"
    ),

    trait = dplyr::recode(
      parts[2],
      ges = "Gestation length",
      litsize = "Litter size",
      lit = "Relative litter mass",
      neo = "Neonate body weight / adult body weight",
      interval = "Interbirth interval"
    ),

    covariates = dplyr::recode(
      parts[3],
      none = "None",
      long = "Longevity",
      adult = "Adult body mass",
      both = "Longevity + adult body mass",
      female = "Female body mass",
      longfemale = "Longevity + female body mass"
    ),

    reference = dplyr::recode(
      parts[5],
      epi = "Epitheliochorial",
      endo = "Endotheliochorial",
      hemo = "Hemochorial",
      vil = "Villous",
      trab = "Trabecular",
      laby = "Labyrinthine"
    )
  )
}

model_metadata <- dplyr::bind_rows(
  lapply(model_names, parse_model_name)
)

# ============================================================
# 6. Build model summary table using rr2::R2_pred
# ============================================================

# Build table first WITHOUT R2
pgls_summary_table <- dplyr::bind_rows(
  lapply(model_names, function(model_name) {

    model <- get(model_name)

    tibble::tibble(
      model_name = model_name,
      AIC = AIC(model),
      BIC = BIC(model),
      Lambda = extract_lambda(model),
      logLik = as.numeric(logLik(model))
    )
  })
) %>%
  dplyr::left_join(model_metadata, by = "model_name") %>%
  dplyr::select(
    model_name,
    trait,
    factor,
    reference,
    covariates,
    AIC,
    BIC,
    Lambda,
    logLik
  )

# Add empty columns
pgls_summary_table$baseline_model <- NA_character_
pgls_summary_table$partial_R2 <- NA_real_

# Compute rr2 R2_pred with progress messages
for (i in seq_len(nrow(pgls_summary_table))) {

  message(
    "\n[", i, "/", nrow(pgls_summary_table), "] ",
    pgls_summary_table$model_name[i]
  )

  this_model <- pgls_summary_table$model_name[i]
  this_trait <- pgls_summary_table$trait[i]
  this_covariates <- pgls_summary_table$covariates[i]
  this_factor <- pgls_summary_table$factor[i]

  baseline_name <- get_baseline_model(
    trait = this_trait,
    covariates = this_covariates
  )

  pgls_summary_table$baseline_model[i] <- baseline_name

  if (this_factor == "None (allometric model)") {
    message("  baseline/allometric row: partial_R2 = NA")
    next
  }

  if (is.na(baseline_name) || !exists(baseline_name)) {
    message("  no baseline model found: partial_R2 = NA")
    next
  }

  model_full <- get(this_model)
  model_base <- get(baseline_name)

  if (nobs(model_full) != nobs(model_base)) {
    message(
      "  different nobs: ",
      nobs(model_full), " vs ", nobs(model_base),
      " — partial_R2 = NA"
    )
    next
  }

  message("  baseline: ", baseline_name)

  r2_value <- tryCatch(
    {
      rr2::R2_pred(model_full, model_base)
    },
    error = function(e) {
      message("  ERROR: ", e$message)
      NA_real_
    }
  )

  pgls_summary_table$partial_R2[i] <- as.numeric(r2_value)

  message("  partial_R2 = ", pgls_summary_table$partial_R2[i])

  # Save progress after every model
  write.csv(
    pgls_summary_table,
    "pgls_summary_table_rr2.csv",
    row.names = FALSE
  )
}

```

```{r pPCA and pPCA PGLS}

``` {r pPCA}

# Build pPCA dataset: drop litter_mass, use relative neonate size
library(dplyr)
library(phytools)

# Build pPCA dataset: drop litter_mass, use relative neonate size
ppca_data <- common_species_new_data %>%
  dplyr::select(
    species,
    gestation_length,
    interbirth_interval,
    litter_size,
    neonate_body_size,
    body_mass_g,
    longevity,
    invasiveness,
    interdigitation
  ) %>%
  dplyr::filter(
    !is.na(gestation_length),
    !is.na(interbirth_interval),
    !is.na(litter_size),
    !is.na(neonate_body_size),
    !is.na(body_mass_g),
    !is.na(longevity),
    gestation_length > 0,
    interbirth_interval > 0,
    litter_size > 0,
    neonate_body_size > 0,
    body_mass_g > 0,
    longevity > 0
  ) %>%
  dplyr::mutate(
    log_neonate_body_size = log(neonate_body_size),
    log_gestation_length = log(gestation_length),
    log_interbirth_interval = log(interbirth_interval),
    log_litter_size = log(litter_size),
    log_body_mass_g = log(body_mass_g),
    log_longevity = log(longevity)
  )

#Check rows N
nrow(ppca_data)

ppca_tree <- ape::keep.tip(
  neotree,
  ppca_data$species
)

#Check if tree tip match with data
ppca_data <- ppca_data %>%
  dplyr::filter(species %in% ppca_tree$tip.label)

ppca_data <- ppca_data[
  match(ppca_tree$tip.label, ppca_data$species),
]

identical(ppca_tree$tip.label, ppca_data$species)

#pPCA matrix generation
ppca_matrix <- ppca_data %>%
  dplyr::select(
    log_gestation_length,
    log_interbirth_interval,
    log_litter_size,
    log_neonate_body_size
  ) %>%
  as.matrix()

rownames(ppca_matrix) <- ppca_data$species

#Now I run the phylogenetic PCA
ppca_result <- phytools::phyl.pca(
  ppca_tree,
  ppca_matrix,
  method = "lambda",
  mode = "corr"
)

#Inspect pPCA
ppca_result$Eval / sum(ppca_result$Eval)
ppca_result$Evec

#Save pPCA scores

ppca_loadings <- as.data.frame(ppca_result$L)
ppca_loadings$trait <- rownames(ppca_loadings)

evals <- diag(ppca_result$Eval)

ppca_variance <- data.frame(
  PC = names(evals),
  eigenvalue = as.numeric(evals),
  proportion_variance = as.numeric(evals / sum(evals)),
  cumulative_variance = cumsum(as.numeric(evals / sum(evals)))
)

ppca_variance

ppca_scores <- as.data.frame(ppca_result$S)
ppca_scores$species <- rownames(ppca_scores)

ppca_scores <- ppca_scores %>%
  dplyr::left_join(
    ppca_data %>%
      dplyr::select(
        species,
        invasiveness,
        interdigitation,
        log_body_mass_g,
        log_longevity
      ),
    by = "species"
  )

colnames(ppca_scores)



#Check for NA
head(ppca_scores)
colSums(is.na(ppca_scores))


#Clean loading labels for plotting
ppca_loadings_plot <- ppca_loadings %>%
  dplyr::mutate(
    trait_label = dplyr::recode(
      trait,
      log_gestation_length = "Gestation length",
      log_interbirth_interval = "Interbirth interval",
      log_litter_size = "Litter size",
      log_neonate_body_size = "Neonate size"
    )
  )


#Create pPCA biplot with trait-loading arrows
arrow_multiplier <- 0.8 * max(
  abs(ppca_scores$PC1),
  abs(ppca_scores$PC2),
  na.rm = TRUE
)

ppca_biplot_interdigitation <- ggplot2::ggplot() +
  ggplot2::geom_point(
    data = ppca_scores,
    ggplot2::aes(
      x = PC1,
      y = PC2,
      color = interdigitation
    ),
    alpha = 0.6,
    size = 2
  ) +
  ggplot2::geom_segment(
    data = ppca_loadings_plot,
    ggplot2::aes(
      x = 0,
      y = 0,
      xend = PC1 * arrow_multiplier,
      yend = PC2 * arrow_multiplier
    ),
    arrow = grid::arrow(length = grid::unit(0.25, "cm")),
    linewidth = 0.9,
    color = "black"
  ) +
  ggplot2::geom_text(
    data = ppca_loadings_plot,
    ggplot2::aes(
      x = PC1 * arrow_multiplier * 1.12,
      y = PC2 * arrow_multiplier * 1.12,
      label = trait_label
    ),
    size = 3.8,
    color = "black"
  ) +
  ggplot2::geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.3
  ) +
  ggplot2::geom_vline(
    xintercept = 0,
    linetype = "dashed",
    linewidth = 0.3
  ) +
  ggplot2::theme_minimal(base_size = 14) +
  ggplot2::labs(
    title = "Phylogenetic PCA biplot of life-history traits",
    x = paste0(
      "pPC1 (",
      round(ppca_variance$proportion_variance[1] * 100, 1),
      "%)"
    ),
    y = paste0(
      "pPC2 (",
      round(ppca_variance$proportion_variance[2] * 100, 1),
      "%)"
    ),
    color = "Interdigitation"
  ) +
  ggplot2::theme(
    panel.grid.minor = ggplot2::element_blank(),
    plot.title = ggplot2::element_text(face = "bold")
  )

print(ppca_biplot_interdigitation)

ggplot2::ggsave(
  "ppca_biplot_interdigitation.png",
  ppca_biplot_interdigitation,
  width = 8,
  height = 6,
  dpi = 300
)

#Same pPCA biplot, colored by invasiveness
ppca_biplot_invasiveness <- ggplot2::ggplot() +
  ggplot2::geom_point(
    data = ppca_scores,
    ggplot2::aes(
      x = PC1,
      y = PC2,
      color = invasiveness
    ),
    alpha = 0.6,
    size = 2
  ) +
  ggplot2::geom_segment(
    data = ppca_loadings_plot,
    ggplot2::aes(
      x = 0,
      y = 0,
      xend = PC1 * arrow_multiplier,
      yend = PC2 * arrow_multiplier
    ),
    arrow = grid::arrow(length = grid::unit(0.25, "cm")),
    linewidth = 0.9,
    color = "black"
  ) +
  ggplot2::geom_text(
    data = ppca_loadings_plot,
    ggplot2::aes(
      x = PC1 * arrow_multiplier * 1.12,
      y = PC2 * arrow_multiplier * 1.12,
      label = trait_label
    ),
    size = 3.8,
    color = "black"
  ) +
  ggplot2::geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.3
  ) +
  ggplot2::geom_vline(
    xintercept = 0,
    linetype = "dashed",
    linewidth = 0.3
  ) +
  ggplot2::theme_minimal(base_size = 14) +
  ggplot2::labs(
    title = "Phylogenetic PCA biplot of life-history traits",
    x = paste0(
      "pPC1 (",
      round(ppca_variance$proportion_variance[1] * 100, 1),
      "%)"
    ),
    y = paste0(
      "pPC2 (",
      round(ppca_variance$proportion_variance[2] * 100, 1),
      "%)"
    ),
    color = "Invasiveness"
  ) +
  ggplot2::theme(
    panel.grid.minor = ggplot2::element_blank(),
    plot.title = ggplot2::element_text(face = "bold")
  )

print(ppca_biplot_invasiveness)

ggplot2::ggsave(
  "ppca_biplot_invasiveness.png",
  ppca_biplot_invasiveness,
  width = 8,
  height = 6,
  dpi = 300
)


#Export pPCA plots
ggplot2::ggsave(
  "ppca_biplot_interdigitation.png",
  ppca_biplot_interdigitation,
  width = 8,
  height = 6,
  dpi = 300
)



#Export pPCA tables

ppca_eigenvalues <- diag(ppca_result$Eval)

ppca_variance <- data.frame(
  PC = paste0("PC", seq_along(ppca_eigenvalues)),
  eigenvalue = as.numeric(ppca_eigenvalues),
  proportion_variance = as.numeric(ppca_eigenvalues / sum(ppca_eigenvalues)),
  cumulative_variance = cumsum(
    as.numeric(ppca_eigenvalues / sum(ppca_eigenvalues))
  )
)

ppca_variance

ppca_loadings <- as.data.frame(ppca_result$Evec) %>%
  tibble::rownames_to_column("trait")

ppca_loadings


write.csv(ppca_loadings, "ppca_loadings.csv", row.names = FALSE)
write.csv(ppca_variance, "ppca_variance_explained.csv", row.names = FALSE)
write.csv(ppca_scores, "ppca_scores.csv", row.names = FALSE)


library(openxlsx)

# Create workbook
ppca_workbook <- openxlsx::createWorkbook()

# Add sheets
openxlsx::addWorksheet(ppca_workbook, "pPCA_loadings")
openxlsx::addWorksheet(ppca_workbook, "pPCA_variance")
openxlsx::addWorksheet(ppca_workbook, "pPCA_scores")

# Write data
openxlsx::writeData(
  ppca_workbook,
  sheet = "pPCA_loadings",
  x = ppca_loadings
)

openxlsx::writeData(
  ppca_workbook,
  sheet = "pPCA_variance",
  x = ppca_variance
)

openxlsx::writeData(
  ppca_workbook,
  sheet = "pPCA_scores",
  x = ppca_scores
)

# Save workbook
openxlsx::saveWorkbook(
  ppca_workbook,
  file = "pPCA_results.xlsx",
  overwrite = TRUE
)



#Now PGLS on pPCA
rownames(ppca_scores) <- ppca_scores$species

se_ppca <- rep(0.01, nrow(ppca_scores))
names(se_ppca) <- rownames(ppca_scores)

inter_ppca_pgls_vil <- pglsSEyPagel(
  PC1 ~ interdigitation,
  data = ppca_scores,
  tree = ppca_tree,
  se = se_ppca,
  method = "ML"
)

invas_ppca_pgls_epi <- pglsSEyPagel(
  PC1 ~ invasiveness,
  data = ppca_scores,
  tree = ppca_tree,
  se = se_ppca,
  method = "ML"
)

inter_ppca_pgls_adult_vil <- pglsSEyPagel(
  PC1 ~ interdigitation + log_body_mass_g,
  data = ppca_scores,
  tree = ppca_tree,
  se = se_ppca,
  method = "ML"
)

invas_ppca_pgls_adult_epi <- pglsSEyPagel(
  PC1 ~ invasiveness + log_body_mass_g,
  data = ppca_scores,
  tree = ppca_tree,
  se = se_ppca,
  method = "ML"
)

inter_ppca_pgls_long_vil <- pglsSEyPagel(
  PC1 ~ interdigitation + log_longevity,
  data = ppca_scores,
  tree = ppca_tree,
  se = se_ppca,
  method = "ML"
)

invas_ppca_pgls_long_epi <- pglsSEyPagel(
  PC1 ~ invasiveness + log_longevity,
  data = ppca_scores,
  tree = ppca_tree,
  se = se_ppca,
  method = "ML"
)

inter_ppca_pgls_both_vil <- pglsSEyPagel(
  PC1 ~ interdigitation + log_longevity + log_body_mass_g,
  data = ppca_scores,
  tree = ppca_tree,
  se = se_ppca,
  method = "ML"
)

invas_ppca_pgls_both_epi <- pglsSEyPagel(
  PC1 ~ invasiveness + log_longevity + log_body_mass_g,
  data = ppca_scores,
  tree = ppca_tree,
  se = se_ppca,
  method = "ML"
)

summary(inter_ppca_pgls_vil)

summary(invas_ppca_pgls_epi)

summary(inter_ppca_pgls_adult_vil)

summary(invas_ppca_pgls_adult_epi)

summary(inter_ppca_pgls_long_vil)

summary(invas_ppca_pgls_long_epi)

summary(inter_ppca_pgls_both_vil)

summary(invas_ppca_pgls_both_epi)


```
